\acknowledgments{
Growing up I dreamed of being a scientist, but I did not even know what a real scientist was . Five years ago, BYU opened its arms to me to help me pursue a career as a scientist. My life has been changed dramatically in ways I could not even imagine years ago . I have so much gratitude to BYU and the Computer Science Department for having me here, as well as gratitude for my family, friends, and colleagues for all their support. I am truly and deeply indebted to so many people that there is no way to acknowledge them all, or even any of them, properly, with the limit of words here.

I would like to express my deepest gratitude to my advisor, Dr.\ Michael Jones. It was my great pleasure to work with him in the Computer Generated Natural Phenomena Lab . I am extremely grateful for his help, support, and encouragement over the last five years. I was impressed by his honest and humble personality as well as his sense of humor. I really appreciate his patience in guiding me in my research area. He has all of my respect as a lifelong best friend and mentor. 

Besides my advisor, I am grateful to my thesis committee: Dr.\ Parris Egbert, Dr.\ Bryan Morse, Dr.\ Eric Mercer, and Dr.\ Sean Warnick. Thank you for your time and efforts in reviewing all my documents and materials. I appreciate all the encouragement, insightful comments, and hard questions. With all these, I am growing up and becoming a real researcher.

Thanks to my colleagues for stimulating discussions and a fun environment. I am especially grateful to Anthony Selino, Seth Holladay, Lanny Lin, Cory Reimschussel, Anthony Hall, Bryce Porter, Tonglaga Bao, and Ontario Britton at BYU. We are good friends learning and growing together.

I would like to thank my parents, Zeyu and Minglu, for their unconditional support. If not for them, there would be no me . Thank you for teaching me to be an honest person in all the good and bad times. Thank you for always supporting me and standing on my side. I would also like to especially thank my grandmas. Both of them have left our world but I know they and their love will be with me no matter whom I am and where I am, forever.

Most of all I would like to thank my husband, Steve, who always gave me love and encouragement. Thank you for your protection and supporting a lovely place for me to stay and be able to concentrate on my study and research. 

BYU accepted me and taught me how to be a good and useful person in our society. I have learned to do scientific research, which is one of the most important parts for why I am here. The world is our campus. I will remember the school's instructions and use them as my guidance in my future career and life. My learning will not stop upon receiving my degree, but will move forward to the bigger world campus. I believe that, all the things I've learned from here, in terms of being a good person in both my professional career and personal life, will benefit me for now and the future. I am so proud that I could have the chance to be stamped by BYU.
}